### 2017-11-29 inst/joss/reviewer_response.md

Review Response
-----------

Thanks @daattali for the thoughtful comments and suggestions. In general the following aspects have been adjusted in view of your comments. The README has been significantly adjusted to better reflect the purpose of the package and more typical use cases. The vignette has also been expanded to discuss the `table_id_inject` function in more detail and the source of the input file is shown. Lastly, the code has been adjusted to help improve debugging (i.e. more informative names internally, `&` changed to `&&` where relevant). More specifics are given below in each of these three areas.

## README

- First 6 bullets are related and bundled here. In general the README has been reworked significantly. The introduction to the package has been clarified and expanded slightly to better discuss what I view is a typical use of the package. Some of this has been borrowed from the paper.md JOSS submission. The trivial example has been replaced with a more typical workflow from Rmd to HTML directly with the functions in the package, this is shown in the section "Simple Example of Package Workflow". In the example, I show the raw input Rmd file, the look of the intermediate markdown table with included CSS ids, and then the final code step to compile the document into an HTML. Finally, a screenshot is included to show the look of the final output. I also attempted to change "table" in a few places to "summary table" to reflect the type of table that I feel should be passed to the `table_id_inject` function.
- Added line to README about installation of devtools package.
- Fixed `library(hightlightHTML)` typo.
- The example with `bgtable.html` was removed as this was a contrived example, primarily for testing. A newer example from Rmd to HTML was created to reflect the most useful use case for the package. This file is called `joss.Rmd`.
- The code should be `update_css`. This typo has been fixed from an older version of the package.
- `update_css = FALSE` is mainly used internally and as such has been removed from the master `highlight_html` function, instead just passed internally to the two `highlight_html_cells` and `highlight_html_text` functions.
- Great suggestion on not overwriting input file. This has been remedied to include a warning if the output file location is not specified and the output is then "<input>_out.html".
- Updated package DESCRIPTION to better reflect package goals and maintain consistency with README documentation.

## Vignette

- The GitHub tagline has been edited as well as the README to better reflect the scope of the package and typical use case.
- More time in vignette is given to the `table_id_inject` function. 
- Similar to the README, the raw input file is shown for the typical example.

## Code

- Included better/more tests. This is an area I have improved on since this package was frozen for JOSS review.
- Internal names were replaced to more descriptive names to help improve debugging in the future. Very useful suggestion.
- Adjusted `rgb2hex` function documentation to reflect named vs unnamed list specification.
- Added error handling to catch double NULL case for `rgb2hex` function.
- Switched code where relevant to `&&` instead of `&`. I wasn't aware of the difference actually. Thanks for pointing this out.
- Attempted to clarify the input for the `table_id_inject` function to: A summary table object, most commonly will be a data.frame, but can also be a count table using the `table` function. Let me know if this is more clear.
- `conditions`: additional examples are given and added documentation within the details section to operations that are allowed and which are allowed for each type of variable. 
- Note: I added a rather simple functionality to the `table_id_inject` function. This function now allows for rounding of numeric columns. A by product of the CSS id insertion is that the columns are converted to text. This makes it more difficult to round numeric values to a specific number of decimal places. This feature has been added as an optional argument to the `table_id_inject` function called `num_digits`. This argument takes a single numeric integer that is passed to the `round` function within R.

### 2017-11-29 inst/examples/joss.Rmd
Update example.
Improved documentation based on JOSS review

```{r symm, echo = FALSE}
library(dplyr)
library(highlightHTML)

chick_tab <- chickwts %>%
  group_by(feed) %>%
  summarise(avg_weight = mean(weight),
            sd_weight = sd(weight))

table_id_inject(chick_tab, id = c('#bggrey', '#bgblack'), 
                conditions = c('> 270', '> 300'),
                variable = list('avg_weight', 'avg_weight'))
```

>>>>

```{r symm, echo = FALSE, results = 'asis', message = FALSE}
library(dplyr)
library(highlightHTML)

chickwts %>%
  group_by(feed) %>%
  summarise(avg_weight = mean(weight),
            sd_weight = sd(weight)) %>%
  mutate(feed = as.character(feed)) %>%
  table_id_inject(id = c('#bggrey', '#bgblack', '#bglightred', '#textblue'), 
                conditions = c('> 270', '> 300', '> 60', '== "horsebean"'),
                variable = list('avg_weight', 'avg_weight', 'sd_weight', 'feed'),
                num_digits = 3) %>%
  knitr::kable(format = 'markdown')
```

### 2017-11-29 readme.Rmd
Improved documentation based on JOSS review

The aim of the `highlightHTML` package is to make it easy for users to customize HTML tables by using a single function that adds an `id` selector to the table cell tag.  The ability to customize the HTML table is limited the functionality of CSS and the user knowledge of CSS.   This package was motivated by latest push of using markdown documents which allow the user to focus on content instead of using something like LaTeX where choices between fonts and styling may take a larger role.  This package is motivated by that in that it allows a seemless way to interact with an HTML file to add table decoration.

## Usage 
Installing the package:
```{r install, eval=FALSE}
library(devtools)
install_github('lebebr01/highlightHTML')
```

This package post processes the html file produced from a markdown document.  Suppose you have a table like the following:

Color Name | Number
----------- | ------------
Blue | 5
Green | 35
Orange | 100
Red | 200

To use this package there are two steps, first indicate which cells you would like to change the appearance using CSS (i.e. background color, text color, etc.) by including the name of the id value to pass that cell or cells.  Using the above table example suppose we want to change the number column to blue and red for numbers less than 10 and greater than 150 respectively:

Color Name | Number
----------- | ------------
Blue | 5 #bgblue
Green | 35
Orange | 100
Red | 200 #bgred

The addition of the *#bgblue* and *#bgred* indicates which cells will be changed.  After turning the markdown document into an html file, this package can now be used to post-process the html file.  The post-processing will add an id value for each cell with the *#bgblue* or *#bgred* and remove those from the table.  This will look something like:

```{r function, eval=FALSE}
library(hightlightHTML)
highlight_html(input = "path/to/file", output = "path/to/saved/file", updateCSS = TRUE, tags = c("Vector of CSS code here"))
```

This function should read in the html file to process.  Then it processes the file add the id tags to the specific cells identified.  Lastly if **updateCSS = TRUE**, then the tags will be added to the CSS of the document, otherwise if **FALSE** that will need to be added manually to the html file.  The file is then save to the output location, if no output location is given, the input file will be overwritten.

>>>>

The aim of the `highlightHTML` package is to add some additional formatting to (R)markdown documents to customize the formatting output of HTML files. The syntax uses HTML `id` selectors and Cascading Style Sheets (CSS) specified by the user to add formatting to the HTML output file. The ability to customize the HTML table is limited the functionality of CSS and the user knowledge of CSS.

The hightlightHTML package fits nicely into the workflow of a reproducible research report as the package can dynamically insert the ids into summary tables with R code. In addition, text can be modified directly with simple syntax. Compilation from rmarkdown or markdown to HTML can be done directly from the package which makes use of the render function from the rmarkdown package. Installation and usage details are shown in more detail below.

## Installation 
Installing the package:
```{r install, eval=FALSE}
install.packages("highlightHTML")
```

or the development version with devtools:
```{r devtools, eval = FALSE}
install.packages("devtools")
devtools::install_github('lebebr01/highlightHTMl')
```

## Usage

As mentioned above, text and summary tables can be formatted using CSS ids. The CSS ids are inserted into text by using braces (i.e. \{) with a CSS id hashtag following the opening brace (e.g. \{#bggrey example\}). In the previous example, the word "example" would be identified with the CSS id #bggrey and could be formated using CSS.

There are two ways to create formatting within a markdown summary table. One is useful if the markdown summary table is created manually (i.e. type by hand). In this scenario you simply need to add the hashtag to the desired cell of the summary table. Perhaps the more common workflow for adding hashtags to a summary table would be dynamically with R code. To do this, the function `table_id_inject` is used to create conditional arguments for adding the hashtags to a markdown summary table. This is shown in more detail with the example below.

## Simple Example of Package Workflow

Below is a simple example of an Rmd document that combines both elements, text markup and inserting CSS ids dynamically into a table using R code and the `table_id_inject` function.

````
---
title: "Rmd to HTML"
author: "Brandon LeBeau"
date: "January 13, 2017"
output: html_document
---

## Simple Markup
The `highlightHTML` packages allows for {#bggrey simple markup} to add formatting to text and tables. Using the `rmarkdown` package and the `table_id_inject` function allows users to easily add markup to Rmd documents and render directly to HTML.

```{r symm, echo = FALSE, results = 'asis', message = FALSE}`r ''`
library(dplyr)
library(highlightHTML)

chickwts %>%
  group_by(feed) %>%
  summarise(avg_weight = mean(weight),
            sd_weight = sd(weight)) %>%
  mutate(feed = as.character(feed)) %>%
  table_id_inject(id = c('#bggrey', '#bgblack', '#bglightred', '#textblue'), 
                conditions = c('> 270', '> 300', '> 60', '== "horsebean"'),
                variable = list('avg_weight', 'avg_weight', 'sd_weight', 'feed'),
                num_digits = 3) %>%
  knitr::kable(format = 'markdown')
```
````

The resulting summary table would look like the following in markdown:
```{r symm, echo = FALSE, results = 'asis', message = FALSE}
library(dplyr)
library(highlightHTML)

chickwts %>%
  group_by(feed) %>%
  summarise(avg_weight = mean(weight),
            sd_weight = sd(weight)) %>%
  mutate(feed = as.character(feed)) %>%
  table_id_inject(id = c('#bggrey', '#bgblack', '#bglightred', '#textblue'), 
                conditions = c('> 270', '> 300', '> 60', '== "horsebean"'),
                variable = list('avg_weight', 'avg_weight', 'sd_weight', 'feed'),
                num_digits = 3) %>%
  knitr::kable(format = 'markdown')
```

It is worth discussing the `table_id_inject` function in more detail here. The function takes two required arguments, `id` and `conditionals` and two optional arguments, `variable` and `num_digits`. The two required arguments are a vector of CSS ids (e.g. #bggrey) and a vector of conditional expressions that are evaluated to identify the location to insert the CSS id. These two arguments must be the same length and the CSS id and conditional expression are matched by location. That is, the first element of each argument are matched, the second element of each, and so on. The optional argument, `variable`, specifies which column(s) of the data to evaluate the conditional expression on. By default this argument is empty meaning that all columns are evaluated. If a CSS id is specific to a specific column(s), this argument specified as a list can be included. Finally, the optional `num_digits` argument is used to round the numeric columns. See `round` for more details.

Finally, the Rmd file itself can be passed to the `highlight_html` function which will automatically compile the input file into an HTML output file. The `highlight_html` function takes three main arguments, the path to the input Rmd file (can also be a markdown or HTML file), the path to save the output HTML file, and the CSS styling to be used for the formatting. Below is the example of processing the simple example shown above.

```{r function, eval=FALSE}
library(highlightHTML)
file <- system.file('examples', 'joss.Rmd', package = 'highlightHTML')

tags <- c('#bgblack {background-color: black; color: white;}',
          '#bggrey {background-color: #d3d3d3;}',
          '#bglightred {background-color: #FF6666;}',
          '#textblue {color: blue}')

highlight_html(input = file, output = tempfile(fileext = ".html"),
               tags = tags, browse = TRUE, 
               render = TRUE)
```

The results HTML output file now looks like the following:

![joss output](inst/examples/joss_output.png)

Finally, the `highlight_html` function has an optional argument called `browse`. This argument is a TRUE/FALSE flag, which defaults to TRUE, to indicate whether the HTML output file should be opened in the default browser upon successful compilation. This can be a good way to view the file to ensure the desired formatting was achieved.

## Vignette
Additional examples can be found in the vignette called, The Basics. This file give more information and examples. The vignette should come with the package when installed from CRAN.

```{r vignette, eval = FALSE}
vignette('the_basics', package = 'highlightHTML')
```

If you are using the developmental version from GitHub, you may need to tell the installer to build the vignette directly:

```{r build_vignette, eval = FALSE}
devtools::install_github('lebebr01/highlightHTML', build_vignettes = TRUE)
```
